---
title: "Bio_tut"
output: html_notebook
author: "Maeva Girerd"
---

```{r}
# install.packages("BiocManager")
# BiocManager::install("Bioconductor")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("ggplot2")
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Programming with R
## addition
```{r}
12/3
```

## Data types
```{r}
num_var <- 10.5         # Numeric variable
char_var <- "Hello"     # Character variable
logical_var <- TRUE     # Logical variable
int_var <- as.integer(5) # Integer variable
```

## Variables and assignement
```{r}
x <- 10
y <- 5
z <- x + y
print(z) 
```
## Introduction to vectors
```{r}
vec1 <- c(1, 2, 3, 4, 5)
vec2 <- c("apple", "banana", "orange")

result <- vec1 + vec1
print(result) 
fruits <- c(vec2, "grape")
print(fruits)
```
## R data structures
```{r}
num_var <- 10.5         
char_var <- "Hello"     
logical_var <- TRUE     
int_var <- as.integer(5)
```

## Data Manipulation in R
```{r}
df <- data.frame(
  Name = c("Alice", "Bob", "Charlie"),
  Age = c(25, 30, 22),
  Score = c(90, 85, 78)
)

subset_df <- df[df$Age > 25, ]
print(subset_df)
```
## Data aggragation and summarization
```{r}
# install.packages("dplyr")
#library(dplyr)
```
```{r}
df <- data.frame(
  Group = c("A", "A", "B", "B", "A", "B"),
  Value = c(10, 20, 15, 25, 30, 35)
)

grouped_df <- df %>% group_by(Group) %>% summarize(Mean_Value = mean(Value))
print(grouped_df)
```

## Data transformation
```{r}
#install.packages("tidyr")
# library(tidyr)
```

```{r}
wide_df <- data.frame(
  ID = c(1, 2),
  Jan = c(100, 150),
  Feb = c(120, 160),
  Mar = c(130, 170)
)

long_df <- pivot_longer(wide_df, cols = -ID, names_to = "Month", values_to = "Value")
print(long_df)
```

# Data visualization in R
## Introduction to ggplot2 for data visualization

```{r}
#install.packages("ggplot2")
# library(ggplot2)
```

```{r}
df <- data.frame(
  Category = c("A", "B", "C"),
  Value = c(10, 20, 15)
)

ggplot(data = df, aes(x = Category, y = Value)) +
  geom_bar(stat = "identity")
```


```{r}
df <- data.frame(
  X = c(1, 2, 3, 4, 5),
  Y = c(10, 20, 15, 25, 30)
)

ggplot(data = df, aes(x = X, y = Y)) +
  geom_point()
```


```{r}
df <- data.frame(
  X = c(1, 2, 3, 4, 5),
  Y = c(10, 20, 15, 25, 30)
)

ggplot(data = df, aes(x = X, y = Y, color = "My Data Points")) +
  geom_point() +
  labs(title = "Scatter Plot Example", x = "X Axis", y = "Y Axis") +
  theme_minimal()
```

```{r}
df <- data.frame(
  Category = rep(c("A", "B"), each = 5),
  Value = c(10, 20, 15, 25, 30)
)

ggplot(data = df, aes(x = Category, y = Value)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Category)
```


# Control structures and functions
## Conditional statements 

```{r}
x <- 10

if (x > 0) {
  print("x is positive.")
} else {
  print("x is non-positive.")
}
```
## Loops

```{r}
for (i in 1:5) {
  print(paste("Iteration:", i))
}
```
## Writing and using functions in R

```{r}
square <- function(x) {
  return(x * x)
}

result <- square(5)
print(result) 
```
# Functional programming with apply family function

```{r}
matrix_data <- matrix(1:9, nrow = 3)

row_means <- apply(matrix_data, 1, mean)
print(row_means)
```


# Data import and export
## Reading data

```{r}
csv_data <- read.csv("~/Documents/Stage_Maeva_Girerd/data/tables/csv/data_Bio_tut.csv")
```

```{r}
csv_data %>% filter(name == "Charles")
csv_data %>% filter(height > 150)
csv_data
```

```{r}
csv_data_height_150 <- csv_data %>% filter(height > 150)
csv_data_height_150
print(dim(csv_data))
csv_data
```

```{r}
# Writing data to CSV file
write.csv(x= csv_data_height_150, "~/Documents/Stage_Maeva_Girerd/results/csv/data_Bio_tut_output.csv")
```

```{r}
# == egalité
5 == 6
# != inegalité
"ciao" != "hello"
```


# Statistical analysis in R
## Descriptive statistics

```{r}
# Descriptive statistics in R
data <- c(10, 15, 20, 25, 30)
mean_value <- mean(data)
median_value <- median(data)
sd_value <- sd(data)
print(mean_value, median_value, sd_value)
```
```{r}
random_data <- rnorm(100, mean = 0, sd = 1)
print(head(random_data))
print(length(random_data))
```

```{r}
summary(csv_data)
```
### T-Test
```{r}
# Creating df with info for males and females
female_df <- csv_data %>% filter(sex == "F")
male_df <- csv_data %>% filter(sex == "M")

female_df
male_df
```
```{r}
height_F <- female_df$height
height_M <- male_df$height

# Perform independent t-test
t_test_result <- t.test(height_F, height_M)
print(t_test_result)
```
p-value = 0.8674
p-value > 0.05, cela signifie qu'il n'y a pas de différence statistiquement significative entre la taille des hommes et des femmes

### corrélation
Female is 1
Male is 0
```{r}
csv_data$sex[csv_data$sex =="F"] <- 1
csv_data$sex[csv_data$sex =="M"] <- 0
csv_data
```

blue is 0
green is 1
brown is 2
For practical reasons (just to see how the test works),
we modified the categorical variables into numeric.
```{r}
csv_data$eye_color[csv_data$eye_color =="blue"] <- 0
csv_data$eye_color[csv_data$eye_color =="green"] <- 1
csv_data$eye_color[csv_data$eye_color =="brown"] <- 2
csv_data
```

```{r}
csv_data_for_correlation <- csv_data %>% select(- name)

csv_data_for_correlation$eye_color <- as.integer(csv_data_for_correlation$eye_color)

csv_data_for_correlation$sex <- as.integer(csv_data_for_correlation$sex)

csv_data_for_correlation

cor(csv_data_for_correlation)
```
test de correlation
```{r}
cor.test(x=csv_data_for_correlation$height, y=csv_data_for_correlation$weight)
```
Entre le poids et la taille il y a une corrélation positive forte (cor = 0.9943661), qui est statistiquement significative (p-vvalue = 4.562e-06 < 0.05)

correlation entre le sexe et la couleur des yeux
correlation entre le sexe et le poids

### regéssion linéaire
```{r}
lm_height_weight <- lm(csv_data_for_correlation$height~csv_data_for_correlation$weight)
summary(lm_height_weight)

plot(csv_data_for_correlation$height~csv_data_for_correlation$weight, data=csv_data_for_correlation) 
abline(lm_height_weight)
```


```{r}
sessionInfo()
```